---
title: "ARIMA 2"
author: "Annie Yang"
date: "10/10/2020"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
---

# 1.0 Libraries

```{r, warning=FALSE}
# Core Tidyverse
library(tidyverse)
library(glue)
library(forcats)

# Time Series
library(timetk)
library(tidyquant)
library(tibbletime)

# Visualization
library(cowplot)

# Preprocessing
library(recipes)

# Sampling / Accuracy
library(rsample)
library(yardstick) 

# Modeling
library(keras)
```

# 2.0 Data

```{r, warning=FALSE}
df <-  read.csv("D:/IIT-master/CS-571-Data Preparation & Analysis-master/project/diabetes_ds/sample_dat.csv", sep=",", header=T, stringsAsFactors = T)

value <- df$IMPUTED

db.date <- paste0(df$YEAR,'-',df$MONTH,'-',df$DAY)
db.time <- paste0(df$HOUR,':',df$MIN,':',df$SEC)


cgm_data <- data.frame(value)
cgm_data$index <-as.POSIXct(paste(db.date, db.time), format="%Y-%m-%d %H:%M:%S")

#Re-ordered
cgm_data <- cgm_data[colnames(cgm_data)[c(2, 1)]]

cgm_data <- as_tbl_time(cgm_data, index = index)


cgm_data
```

# 3.0 Exploratory Data Analysis

# 3.1 Visualizing Sunspot Data With Cowplot
```{r, fig.width=16, warning=FALSE, echo=FALSE, fig.height=12}
p1 <- cgm_data %>%
    filter_time("2019-08-01 00:00" ~ "2019-08-31 00:00") %>%
    ggplot(aes(index, value)) +
    geom_point(color = palette_light()[[1]], alpha = 5, size=1) +
    theme_tq() +
    labs(
        title = "From Aug 1st to Aug 31st"
    )
p2 <- cgm_data %>%
    filter_time("2019-08-05 00:00" ~ "2019-08-9 00:00") %>%
    ggplot(aes(index, value)) +
    geom_line(color = palette_light()[[1]], alpha = 0.5) +
    geom_point(color = palette_light()[[1]]) +
    geom_smooth(method = "loess", span = 0.2, se = FALSE) +
    theme_tq() +
    labs(
        title = "From Aug 5 to Aug 9nd (4 days data)"
    )

p_title <- ggdraw() + 
    draw_label("CGM dataset 2019", size = 18, fontface = "bold", colour = palette_light()[[1]])
plot_grid(p_title, p1, p2, ncol = 1, rel_heights = c(0.1, 1, 1))
```


```{r, warning=FALSE, echo=FALSE}

cgm <- cgm_data %>%
    filter_time("2019-08-05 00:00" ~ "2019-08-09 00:00")

periods_train <- 48 # 1 week
periods_test  <- 8 # 2 hours
skip_span     <- 56

rolling_origin_resamples <- rolling_origin(
    cgm,
    initial    = periods_train,
    assess     = periods_test,
    cumulative = FALSE,
    skip       = skip_span
)

rolling_origin_resamples
```
```{r}
# Plotting function for a single split
plot_split <- function(split, expand_y_axis = TRUE, alpha = 1, size = 1, base_size = 14) {
    
    # Manipulate data
    train_tbl <- training(split) %>%
        add_column(key = "training") 
    
    test_tbl  <- testing(split) %>%
        add_column(key = "testing") 
    
    data_manipulated <- bind_rows(train_tbl, test_tbl) %>%
        as_tbl_time(index = index) %>%
        mutate(key = fct_relevel(key, "training", "testing"))
    
    # Collect attributes
    train_time_summary <- train_tbl %>%
        tk_index() %>%
        tk_get_timeseries_summary()
    
    test_time_summary <- test_tbl %>%
        tk_index() %>%
        tk_get_timeseries_summary()
    
    # Visualize
    g <- data_manipulated %>%
        ggplot(aes(x = index, y = value, color = key)) +
        geom_line(size = size, alpha = alpha) +
        theme_tq(base_size = base_size) +
        scale_color_tq() +
        labs(
            title    = glue("Split: {split$id}"),
            subtitle = glue("{train_time_summary$start} to {test_time_summary$end}"),
            y = "", x = ""
        ) +
        theme(legend.position = "none") 
    
    
    
    return(g)
}
```

```{r, warning=FALSE}
rolling_origin_resamples$splits[[1]] %>%
    plot_split() +
    theme(legend.position = "bottom")
```

```{r}
# Plotting function that scales to all splits 
plot_sampling_plan <- function(sampling_tbl, expand_y_axis = TRUE, 
                               ncol = 4, alpha = 1, size = 1, base_size = 14, 
                               title = "Sampling Plan") {
    
    # Map plot_split() to sampling_tbl
    sampling_tbl_with_plots <- sampling_tbl %>%
        mutate(gg_plots = map(splits, plot_split, 
                              expand_y_axis = expand_y_axis,
                              alpha = alpha, base_size = base_size))
    
    # Make plots with cowplot
    plot_list <- sampling_tbl_with_plots$gg_plots 
    
    p_temp <- plot_list[[1]] + theme(legend.position = "bottom")
    legend <- get_legend(p_temp)
    
    p_body  <- plot_grid(plotlist = plot_list, ncol = ncol)
    
    p_title <- ggdraw() + 
        draw_label(title, size = 18, fontface = "bold", colour = palette_light()[[1]])
    
    g <- plot_grid(p_title, p_body, legend, ncol = 1, rel_heights = c(0.05, 1, 0.05))
    
    return(g)
    
}
```


```{r, fig.width=12, fig.height=10}
rolling_origin_resamples %>%
    plot_sampling_plan(expand_y_axis = T, ncol = 4, alpha = 1, size = 1, base_size = 10, 
                       title = "Plot of Train/Test Splits")
```

```{r}
set <- list()

for (i in 1:20) {
  split    <- rolling_origin_resamples$splits[[i]]
  split_id <- rolling_origin_resamples$id[[i]]
  
  df_trn <- training(split)
  df_tst <- testing(split)
 
  set[[i]] <- bind_rows(
    df_trn %>% add_column(key = "training"),
    df_tst %>% add_column(key = "testing")
    ) %>%
    as_tbl_time(index = index)
}
```

```{r, fig.width=12, fig.height=24, warning=FALSE}
library(tseries)
library(forecast)

ts_trn <- list()
ts_tes <- list()

for (i in 1:20) {
  #split train/test
  ts <- ts(set[[i]]$value)
  ts_trn[[i]] <- head(ts, n=48)
  h <- length(ts) - length(ts_trn[[i]])
  ts_tes[[i]] <- tail(ts, h)
}

```


Dicky-Fuller Test
```{r}
library(tseries)
ts <- ts_trn[[20]]
adf.test(ts)
adf.test(diff(ts, lag=1))
adf.test(diff(ts, lag=2))
adf.test(diff(ts, lag=3))
adf.test(diff(ts, lag=4))
adf.test(diff(ts, lag=5))
```

```{r}
#kpss.test(ts, null = c("Level", "Trend"), lshort = TRUE)
kpss.test(ts, null = c("Level", "Trend"), lshort = TRUE)
kpss.test(diff(ts, lag=1), null = c("Level", "Trend"), lshort = TRUE)
kpss.test(diff(ts, lag=2), null = c("Level", "Trend"), lshort = TRUE)
kpss.test(diff(ts, lag=3), null = c("Level", "Trend"), lshort = TRUE)
kpss.test(diff(ts, lag=4), null = c("Level", "Trend"), lshort = TRUE)
kpss.test(diff(ts, lag=5), null = c("Level", "Trend"), lshort = TRUE)
```
```{r}
Box.test(ts, type="Ljung-Box", lag=1)
Box.test(diff(ts, lag=1), type="Ljung-Box", lag=1)
Box.test(diff(ts, lag=2), type="Ljung-Box", lag=1)
Box.test(diff(ts, lag=3), type="Ljung-Box", lag=1)
```
ACF/PACF
```{r, fig.width=12, fig.height=4}
for (i in 20) {
  
  diiff.ts = diff(ts_trn[[i]], lag=1)
  
  par(mfrow=c(1, 2))
  
  pacf(diiff.ts,lag=20)
  acf(diiff.ts,lag=50)
  
  #adf.test(ts)
}

```
# Our PACF slowly tapers to 0, although it has two spikes at lags 1 and 2. On the other side, our ACF shows a tapering pattern, with lags slowly degrading towards 0. The table below can be used to help identify patterns, and what model conclusions we can make about those patterns. 
# Since the ACF plot apers to 0 in some fashion and the PACF pattern has non-zero values at first p points; then lead to zero values elsewhere, we can conclude it as a AR(p) Model.
```
decompose(diiff.ts)
```

```{r}
arimaModel=arima(ts_trn[[20]], order=c(3, 1, 0))
print(arimaModel)
```

```{r}
qqnorm(arimaModel$residuals, main = "Normal Q-Q Plot")
```

```{r}
fores=forecast(arimaModel, level = c(95), length(ts_tes[[20]]))
print(fores)
```

```{r}
#pre.fores=predict(arimaModel, level = c(95), length(ts_tes[[20]]))
print(pre.fores)
length(ts_tes[[20]])
```

```{r}
#library(ggfortify)
#res <- rbind(ts_tes[[20]], fores$fitted)
autoplot(fores, predict.colour = 'red1', ts.colour = 'gray23',predict.linetype = 'dashed', size=1, predict.size = 1.1) 
```

```
library(ggfortify)
autoplot(fores, predict.colour = 'red1', ts.colour = 'gray23',predict.linetype = 'dashed', size=1, predict.size = 1.1) 
```

```{r}
getPerformance = function(pred, val) {
    res = pred - val
    MAE = sum(abs(res))/length(val)
    RSS = sum(res^2)
    MSE = RSS/length(val)
    RMSE = sqrt(MSE)
    perf = data.frame(MAE, RSS, MSE, RMSE)
}

library(DMwR)
#length(fores$fitted)

#accmeasures=regr.eval(ts_tes[[1]], fores$fitted)
#accmeasures

my.accmeasures=getPerformance((ts_tes[[20]]), matrix(tail(fores$fitted,n=length(ts_tes[[20]]))))
my.accmeasures

```


```{r}
AutoArimaMod = auto.arima(ts_trn[[20]])
print(AutoArimaMod)
```
```{r}
qqnorm(AutoArimaMod$residuals, main = "Normal Q-Q Plot")
```

```{r}
auto.fores=forecast(AutoArimaMod, level = c(95), length(ts_tes[[20]]))
print(auto.fores)
```

```{r}
library(ggfortify)

autoplot(auto.fores, predict.colour = 'red1', ts.colour = 'gray23',predict.linetype = 'dashed', size=1, predict.size = 1.1) 
```

```{r}
auto.accmeasures=getPerformance((ts_tes[[20]]), matrix(tail(auto.fores$fitted,n=length(ts_tes[[20]]))))
auto.accmeasures
```

```{r}
arima_model <- list()
for (i in 1:20) {
  ts <- ts_trn[[i]]
  
  arima_model[[i]] = auto.arima(ts)
  
  #show(model[[i]])
}
```

```{r}
arima_model[[20]]
```

```{r, warning=FALSE}
library(ggfortify)

preds <- list()
res <- list()

for (i in 1:20) {
  ts <- ts_tes[[i]]
  
  preds[[i]] <- forecast(arima_model[[i]], level = c(95), length(ts))
  
  res[[i]] <- preds[[i]]$fitted
  
  #res_plot <- rbind(ts, res[[i]])
  #res_plts[i] <- autoplot(preds[[i]], predict.colour = 'red1', ts.colour = 'gray23',predict.linetype = 'dashed', size=1, predict.size = 1.1)
}
```

```{r,fig.width=12}
# Plotting function to all splits 
plot_resulting_plan <- function(preds,
                               ncol = 4, alpha = 1, size = 1, base_size = 14,
                               predict.colour = 'red1', ts.colour = 'gray23',predict.linetype = 'dashed', predict.size = 1.1,
                               title = "Forecast Result plots") {
    
    # Map plot_split() to sampling_tbl
    gg_plots <- map(preds, autoplot, predict.colour = predict.colour, ts.colour = ts.colour,predict.linetype = predict.linetype, size=size, predict.size = predict.size)
    
    # Make plots with cowplot
    plot_list <- gg_plots 
    
    p_temp <- plot_list[[1]] + theme(legend.position = "bottom")
    legend <- get_legend(p_temp)
    
    p_body  <- plot_grid(plotlist = plot_list, ncol = ncol)
    
    p_title <- ggdraw() + 
        draw_label(title, size = 18, fontface = "bold", colour = palette_light()[[1]])
    
    g <- plot_grid(p_title, p_body, legend, ncol = 1, rel_heights = c(0.05, 1, 0.05))
    
    return(g)
    
}
```


```{r, fig.width=12, fig.height=9}
plot_resulting_plan(preds=preds, title = "Forecast Result plots", alpha = 0.1)
```

```{r}
getPerformance = function(pred, val) {
    res = pred - val
    MAE = sum(abs(res))/length(val)
    RSS = sum(res^2)
    MSE = RSS/length(val)
    RMSE = sqrt(MSE)
    perf = data.frame(MAE, RSS, MSE, RMSE)
}
```


```{r}
res_perf <- list()
for (i in 1:length(preds)) {
  res_perf[[i]] <- getPerformance(matrix(ts_tes[[i]]), matrix(tail(res[[i]], n=length(ts_tes[[i]]))))
}
```

```{r}
res_perf
```

```{r}
mae.val = 0
rss.val = 0
mse.val = 0
rmse.val = 0

for (i in 1:20) {
  mae.val = mae.val+res_perf[[i]][,1]
  rss.val = rss.val+res_perf[[i]][,2]
  mse.val = mse.val+res_perf[[i]][,3]
  rmse.val = rmse.val+res_perf[[i]][,4]
}
mae.val
rss.val
mse.val
rmse.val
avg.perf = data.frame(mae.val/20, rss.val/20, mse.val/20, rmse.val/20)
show(avg.perf)
```























