---
title: "Time series - ARIMA model"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
author: Annie Yang
---

# `Load` the data set (*.csv file) & Extract the `IMPUTED` column.
```{r}
df <-  read.csv("D:/IIT-master/CS-571-Data Preparation & Analysis-master/project/diabetes_ds/sample_dat.csv", sep=",", header=T, stringsAsFactors = T)

glu_level<- df$IMPUTED
diabetes.df <- data.frame(glu_level)
sum(is.na(diabetes.df))
#View(diabetes.df)
```

Have `524` missing values.

# Combine columns to form the `timestamp` as the `Date Object`.
```{r}
db.date <- paste0(df$YEAR,'-',df$MONTH,'-',df$DAY)
db.time <- paste0(df$HOUR,':',df$MIN,':',df$SEC)

diabetes.df$date_time <- as.POSIXct(paste(db.date, db.time), format="%Y-%m-%d %H:%M:%S")

#diabetes.df$Date <- as.Date(db.date)
#Re-ordered
diabetes.df <- diabetes.df[colnames(diabetes.df)[c(2, 1)]]

#head(diabetes.df)
#tail(diabetes.df)
```

## Take substes (by Month)
```{r}
sub.dat = subset(diabetes.df, date_time >= as.POSIXct("2019-08-01 00:00") & date_time < as.POSIXct("2019-08-02 00:00"))
#sum(is.na(ts.84))
#ts.Feb

```

```{r}
t = sub.dat$glu_level
y_stationary <- rnorm(length(t),mean=1,sd=1) # the stationary time series (ts)
y_trend      <- cumsum(rnorm(length(t),mean=1,sd=4))+t/100 # our ts with a trend
# lets normalize each for simplicity
y_stationary<- y_stationary/max(y_stationary) 
y_trend      <- y_trend/max(y_trend) 
```

### Ljung-Box test for independence
Quantitatively, we can also use built-in test for testing stationariy. First, the Ljung-Box test examines whether there is significant evidence for non-zero correlations at given lags (1-25 shown below), with the null hypothesis of independence in a given time series (a non-stationary signal will have a low p-value).
```{r}
lag.length = 25
Box.test(y_stationary, lag=lag.length, type="Ljung-Box") # test stationary signal
```

```{r}
Box.test(y_trend,      lag=lag.length, type="Ljung-Box") # test nonstationary signal
```

# Augmented Dickey–Fuller (ADF) t-statistic test for unit root
Another test we can conduct is the Augmented Dickey–Fuller (ADF) t-statistic test to find if the series has a unit root (a series with a trend line will have a unit root and result in a large p-value).

```{r, warning=FALSE}
options(warn=-1)
library(tseries)

adf.test(y_stationary)
```

```{r}
adf.test(y_trend)
```


# Kwiatkowski-Phillips-Schmidt-Shin (KPSS) for level or trend stationarity

Lastly, we can test if the time series is level or trend stationary using the Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test. Here we will test the null hypothesis of trend stationarity (a low p-value will indicate a signal that is not trend stationary, has a unit root):

```{r, warning=FALSE}
kpss.test(y_stationary, null="Trend")
```

```{r, warning=FALSE}
kpss.test(y_trend, null="Trend")
```

```{r}
ts <- ts(sub.dat$glu_level)
print(ts)
options(warn=-1)
library(tseries)

adf.test(ts)
```

Split into training/test set (80/20) & Plot
```{r, fig.width=10, fig.height=5}
library(forecast)
#train.idx <- sample(1:nrow(sub.dat), size=0.8*nrow(sub.dat))


#dat.train <- sub.dat[train.idx, ]
#dat.test <- sub.dat[-train.idx, ]

ts.train <- head(ts, round(length(ts) * 0.95))
h <- length(ts) - length(ts.train)
ts.test <- tail(ts, h)
length(ts)
length(ts.train)
length(ts.test)

#plotting the data
autoplot(ts.train, size = 2) + autolayer(ts.test, size = 2)
```

ACF/PACF
```{r}
pacf(diff(ts.train, lag=2), lag=20)
acf(diff(ts.train, lag=2), lag=20)
```

Apply auto.arima() function
```{r, warning=FALSE}
library(forecast)
model = auto.arima(ts.train)
model
```

Forcast
```{r, fig.width=10, fig.height=5}
#plot(forecast(model,length(ts.test)))
bg.preds <- forecast(model, level = c(95), length(ts.test))
#print(bg.preds$fitted)
#autoplot(bg.preds, size=1.1, predict.size = 1.1, predict.colour = 'red')
#rbind(ts.train, bg.preds)
#lines(ts,col="blue")
```

```{r}
length(bg.preds$fitted)
```

```{r, warning=FALSE, fig.width=10, fig.height=5}
library(ggfortify)
res <- rbind(ts, bg.preds$fitted)
autoplot(bg.preds, predict.colour = 'red1', ts.colour = 'gray23',predict.linetype = 'dashed', size=1, predict.size = 1.1) 
```

































































