---
title: "Multivariate processing and exploration"
output: html_notebook
author: Emily Tao
---

Loading Data

```{r}
df=read.table('E:/Documents/iit-coursework/csp-571/S_SET03_CASE005_M2_P2.csv',
               header=TRUE, sep=',', na.strings='NaN')
cgm_string='imputed' #set imputed/denoised as CGM value
if (cgm_string=='imputed') { 
  cgm=df$IMPUTED
} else { 
  cgm=df$Denoised
}
```

Converting time information to timestamp objects (from Annie)

```{r}
db.date <- paste0(df$YEAR,'-',df$MONTH,'-',df$DAY)
db.time <- paste0(df$HOUR,':',df$MIN,':',df$SEC)
date_time <- as.POSIXct(paste(db.date, db.time), format="%Y-%m-%d %H:%M:%S")
```

Assembling the cleaned data frame

```{r}
diabetes.df <- data.frame(date_time,cgm)
```

Function to fill NA values with the previous value (or next value, if there is no previous value as in the start of the data file)
```{r}
fill.withprev = function(x)
{
    first.ind = which(!is.na(x))[1]
    prev=x[first.ind]
    for (i in 1:length(x))
    {
      if(is.na(x[i]))
        x[i]=prev
      else
        prev=x[i]
    }
    return(x)
}
  
```

Handling NA values in sparse variables

```{r}
bolus=df$bolus
bolus[is.na(bolus)]=0

basal=df$BASAL

library(mltools)
library(data.table)
activity_names=df$Activity_Name
activity_names[is.na(activity_names)]=0
activity_names=factor(activity_names)
activity_oneh <- one_hot(as.data.table(activity_names)) 

distance_value=df$Distance_value
distance_value[is.na(distance_value)]=0

activity_duration=df$Activity_Duration
activity_duration[is.na(activity_duration)]=0

insulin_onboard=fill.withprev(df$insulinOnBoard)

insulin_sensitivity=fill.withprev(df$insulinSensitivity)

smbg_new=df$smbg_new
smbg_new[is.na(smbg_new)]=0

carb_input_new=df$carb_input_new
carb_input_new[is.na(carb_input_new)]=0
```

Adding sparse variables to the dataframe

```{r}
diabetes.df$bolus=bolus
diabetes.df$basal=basal
diabetes.df$distance_value=distance_value
diabetes.df$activity_duration=activity_duration
diabetes.df=cbind(diabetes.df,activity_oneh)
diabetes.df$insulin_onboard=insulin_onboard
diabetes.df$insulin_sensitivity=insulin_sensitivity
diabetes.df$smbg_new=smbg_new
diabetes.df$carb_input_new=carb_input_new
```

## Data Exploration

Set the time window

```{r}
#Set the time window
start_time = as.POSIXct("2019-05-01 00:00")
end_time = as.POSIXct("2019-05-03 00:00")
window = which(diabetes.df$date_time >= start_time & diabetes.df$date_time < end_time)
```

Plot some features overlaid on one plot

```{r fig.height=6, fig.width=12}
#Draw the plot
plot(diabetes.df$cgm[window],type='l',main='First 48 hours of IMPUTED cgm values',ylim=c(0,16))
lines(diabetes.df$basal[window],col='green')
lines(diabetes.df$bolus,col='blue')
lines(diabetes.df$distance_value,col='red')
lines(diabetes.df$activity_names_1,col='purple')
ad_norm = diabetes.df$activity_duration/max(diabetes.df$activity_duration)
lines(ad_norm*15,col='orange')
legend('topright',inset=0.01, legend=c('CGM','basal insulin','bolus insulin','distance traveled','activity duration','activity name'),col=c('black','green','blue','red','orange','purple'), lty=1)
```

Plot everything in their own window

```{r fig.height=6, fig.width=10}
#Draw the plot
par(mfcol=c(5,2))
par('mar'=c(2,2,2,2))
colors = rainbow(9)
names = colnames(diabetes.df)[c(3:6,8,11:14)]
plot(diabetes.df$cgm[window],type='l',lwd=2,main='First 48 hours of IMPUTED cgm values',ylim=c(0,16))
legend('topright',inset=0.01, legend=c('cgm'),col=c('black'),lty=1)
plot(diabetes.df$bolus[window],type='l',lwd=2,col=colors[1])
legend('topright',inset=0.01, legend=c(names[1]),col=colors[1],lty=1)
plot(diabetes.df$basal[window],type='l',lwd=2,col=colors[2])
legend('topright',inset=0.01, legend=c(names[2]),col=colors[2],lty=1)
plot(diabetes.df$distance_value[window],type='l',lwd=2,col=colors[3])
legend('topright',inset=0.01, legend=c(names[3]),col=colors[3],lty=1)
plot(diabetes.df$activity_duration[window],type='l',lwd=2,col=colors[4])
legend('topright',inset=0.01, legend=c(names[4]),col=colors[4],lty=1)
plot(diabetes.df$activity_names_1[window],type='l',lwd=2,col=colors[5])
legend('topright',inset=0.01, legend=c(names[5]),col=colors[5],lty=1)
plot(diabetes.df$insulin_onboard[window],type='l',lwd=2,col=colors[6])
legend('topright',inset=0.01, legend=c(names[6]),col=colors[6],lty=1)
plot(diabetes.df$insulin_sensitivity[window],type='l',lwd=2,col=colors[7])
legend('topright',inset=0.01, legend=c(names[7]),col=colors[7],lty=1)
plot(diabetes.df$smbg_new[window],type='l',lwd=2,col=colors[8])
legend('topright',inset=0.01, legend=c(names[8]),col=colors[8],lty=1)
plot(diabetes.df$carb_input_new[window],type='l',lwd=2,col=colors[9])
legend('topright',inset=0.01, legend=c(names[9]),col=colors[9],lty=1)
```

Plot correlations between features.

```{r}
library(corrplot)
remna.ind=which(!is.na(diabetes.df$cgm))
corrs=cor(diabetes.df[remna.ind,2:ncol(diabetes.df)])
corrplot(corrs)
```

Assessing where are the missing CGM sections

```{r}
nainds=which(is.na(diabetes.df$cgm))
starts_lagged=which(diff(nainds)>1)
starts=c(nainds[1],nainds[starts_lagged+1])
ends=c(nainds[starts_lagged],nainds[length(nainds)])
naseq=matrix(c(starts,ends),nrow=length(starts))
sprintf('The NA start and end points are: ')
naseq
diabetes.df$date_time[naseq]
```
save the data frame
```{r}
save(diabetes.df,file=paste0("diabetes_mvar_",cgm_string,".Rda"))
```

