---
title: "Multivariate LSTM with xts package"
output: html_notebook
author: Emily Tao
---

Load saved dataframe

```{r}
#load("diabetes_mvar_denoised.Rda")
#load("diabetes_mvar_imputed.Rda")
load("diabetes_mvar_imputed_complete.Rda")

usecols=c('cgm',
          'bolus',
          'basal',
          'distance_value',
          'activity_duration',
          'activity_names_1',
          'activity_names_11',
          'insulin_onboard')
date_time=diabetes.df$date_time
diabetes.df=diabetes.df[,usecols]
```

Create xts object

```{r}
library(xts)
diabetes.xts=xts(diabetes.df,order.by=date_time)
head(diabetes.xts)
```
plotting
```{r}
plot(diabetes.xts['2019-05-21/2019-05-22'],major.ticks='days',minor.ticks='hours',main=NULL,ylim=c(0,20))
```

Diff the cgm values
```{r}
usedif=T
diabetes.dif=diabetes.xts

if (usedif)
{
  cgm.diff = diff(diabetes.xts[,'cgm'],differences=1,na.pad=T)
  diabetes.dif[,'cgm'] = cgm.diff
}
```

Set the time window (Exclude NA sections) - old method, ignore
```{r}
#window=window[1:(8000+13)]
#diabetes.win = diabetes.dif['2019-05-21/2019-06-13']
#diabetes.win.nodif = diabetes.xts['2019-05-21/2019-06-13']
#which(is.na(diabetes.win[,'cgm'])) #check NA values
```

Rescaling/normalization functions
```{r}
#function 'NORMALIZE':
#   take x=data, params=c(newmean,newsd)
#   return rescaled_x, c(originalmean,originalsd)
normalize = function(x, params=c(0,1),reverse=F)
{
  newmean=params[1]
  newsd=params[2]
  xbar=ifelse(reverse,0,mean(x, na.rm=T))
  s=ifelse(reverse,1,sd(x, na.rm=T))
  scaled_x = (x-xbar)/s*newsd+newmean
  return(list(x=scaled_x,param_orig=c(xbar,s)))
}

#function 'RESCALE':
#   take x=data, params=c(newmin,newmax)
#   return rescaled_x, c(originalmin,originalmax)
rescale = function(x, params=c(0,1))
{
  newmin=params[1]
  newmax=params[2]
  oldmin=min(x, na.rm=T)
  oldmax=max(x, na.rm=T)
  scaled_x = (x-oldmin)/(oldmax-oldmin)*(newmax-newmin)+newmin
  return(list(x=scaled_x,param_orig=c(oldmin,oldmax)))
}
```

Rescaling/normalization. 
- Normalized to mean=0, sd=1: {cgm}
- Rescaled to min=0, max=1:  {basal, bolus, distance_value}
```{r}
#diabetes.res=diabetes.win #from old windowing method
diabetes.res=diabetes.dif
orig.pars=list()
for (feat in usecols)
{
  if (feat=='cgm' || feat=='basal')
    rescaled=normalize(diabetes.res[,feat],c(0,1))
  else
    rescaled=rescale(diabetes.res[,feat],c(0,2))
  rescaled.feat=rescaled$x
  orig.par=rescaled$param_orig
  diabetes.res[,feat]=rescaled.feat
  orig.pars=c(orig.pars,list(orig.par))
}
plot(diabetes.res['2019-05-21/2019-05-22'],major.ticks='days',minor.ticks='hours',main=NULL)
```
```{r}
library(corrplot)
corrs=cor(diabetes.res)
corrplot(corrs)
```

Detect NA runs - current sampling method samples from the whole data file, and removes any batches with NA values.

```{r}
detect_na = function(x)
{
  if(any(is.na(x)))
    return(TRUE)
  else
    return(FALSE)
}

detect_na_inds = function(x)
{
  inds=logical(nrow(x))
  for (i in 1:nrow(x))
    inds[i]=detect_na(x[i,,])
  return(inds)
}
```

Reshape to [batches, timesteps, features]

```{r}
timesteps=12*36+48
n_batches=100
diabetes.lag = lag(diabetes.res,timesteps:0)
diabetes.mat.full = as.matrix(diabetes.lag)
dim(diabetes.mat.full) = c(nrow(diabetes.mat.full),length(usecols),timesteps+1)
diabetes.mat.full=aperm(diabetes.mat.full,c(1,3,2))

#from old windowing method
#diabetes.mat=diabetes.mat[seq(from=timesteps+1,to=nrow(diabetes.lag),length.out=1000),,] 

na_inds = which(detect_na_inds(diabetes.mat.full))
diabetes.mat = diabetes.mat.full[-na_inds,,]
sample=seq(1,nrow(diabetes.mat),length.out=n_batches)
diabetes.mat=diabetes.mat[sample,,]
```

Test-train split
```{r}
npred=48
n=nrow(diabetes.mat)*0.75
N=nrow(diabetes.mat)
y_train = diabetes.mat[1:n,(ncol(diabetes.mat)-npred+1):ncol(diabetes.mat),1]
x_train = diabetes.mat[1:n,1:(ncol(diabetes.mat)-npred),]
y_test = diabetes.mat[(n+1):N,(ncol(diabetes.mat)-npred+1):ncol(diabetes.mat),1]
x_test = diabetes.mat[(n+1):N,1:(ncol(diabetes.mat)-npred),]
```

create matrix of original-scaled y values
```{r}
diabetes.lag.nodif = lag(diabetes.xts[,'cgm'],timesteps:0)
diabetes.mat.nodif = as.matrix(diabetes.lag.nodif)
diabetes.mat.nodif = diabetes.mat.nodif[-na_inds,] 
diabetes.mat.nodif = diabetes.mat.nodif[sample,]
y_test.nodif = diabetes.mat.nodif[(n+1):N,(ncol(diabetes.mat)-npred+1):ncol(diabetes.mat)]
```

Some parameters
```{r}
# specify required arguments
X_shape2 = dim(x_train)[2]
X_shape3 = dim(x_train)[3]
batch_size = 1                # must be a common factor of both the train and test samples
units = 30
```

Create LSTM model
```{r}
library(keras)
library(tensorflow)
model <- keras_model_sequential() 
model%>%
  layer_lstm(units, 
             batch_input_shape = c(batch_size, X_shape2, X_shape3),
             return_sequences = F, 
             stateful= TRUE)%>%
  layer_dense(units = 50) %>%
  layer_dense(units = npred)
model
```
Compile model 
```{r}
model %>% compile(
  loss = 'mean_absolute_error',
  optimizer = optimizer_adam( lr= 0.05, decay = 1e-6 ),  
  metrics = c('mean_squared_error')
)
```

Train
```{r}
Epochs = 15
for(i in 1:Epochs ){
  model %>% fit(x_train, y_train, 
                epochs=1, 
                batch_size=batch_size, 
                verbose=1, 
                shuffle=FALSE)
  model %>% reset_states()
}
```


Predict on test sections
```{r}
L = nrow(x_test)/batch_size
predictions = numeric(nrow(x_test)*npred)
predictions.raw = numeric(nrow(x_test)*npred)
dim(predictions)=c(L,npred)
dim(predictions.raw)=c(L,npred)
cgm=diabetes.mat.nodif#as.matrix(diabetes.xts[,'cgm'])
for(i in 1:L){
     X = x_test[(1+(i-1)*batch_size):(i*batch_size),,]
     dim(X) = c(batch_size,X_shape2,X_shape3)
     yhat.raw = model %>% predict(X, batch_size=batch_size)
     # invert scaling
     yhat = normalize(yhat.raw, orig.pars[[1]], reverse=T)$x
     # invert differencing
     if (usedif) {
       yhat[1]=yhat[1]+cgm[i,(timesteps-npred-1)]
       for (j in 2:npred)
           yhat[j]=yhat[j-1]+yhat[j]
     }
     # store
     predictions[i,] <- yhat
     predictions.raw[i,] <- yhat.raw
}
```

Predict if user took bolus/activity right now (just for fun)
```{r}
testbol = mean(diabetes.res$bolus[diabetes.res$bolus>0])*2
x_test.addbol = x_test
x_test.addbol[,ncol(x_test.addbol),2] = x_test.addbol[,ncol(x_test.addbol),2] + testbol
testact = mean(diabetes.res$distance_value[diabetes.res$distance_value>0])
x_test.addact = x_test
x_test.addact[,ncol(x_test.addact),4] = x_test.addact[,ncol(x_test.addact),4] + testbol

L = nrow(x_test)/batch_size
predictions.addbol = numeric(nrow(x_test)*npred)
predictions.addact = numeric(nrow(x_test)*npred)
dim(predictions.addbol)=c(L,npred)
dim(predictions.addact)=c(L,npred)
for(i in 1:L){
     X = x_test.addbol[(1+(i-1)*batch_size):(i*batch_size),,]
     dim(X) = c(batch_size,X_shape2,X_shape3)
     yhat = model %>% predict(X, batch_size=batch_size)
     # invert scaling
     yhat = normalize(yhat, orig.pars[[1]], reverse=T)$x
     # invert differencing
     yhat[1]=yhat[1]+cgm[i,(timesteps-npred-1)]
     for (j in 2:npred)
       yhat[j]=yhat[j-1]+yhat[j]
     # store
     predictions.addbol[i,] <- yhat
     
     X = x_test.addact[(1+(i-1)*batch_size):(i*batch_size),,]
     dim(X) = c(batch_size,X_shape2,X_shape3)
     yhat = model %>% predict(X, batch_size=batch_size)
     # invert scaling
     yhat = normalize(yhat, orig.pars[[1]], reverse=T)$x
     # invert differencing
     yhat[1]=yhat[1]+cgm[i,(timesteps-npred-1)]
     for (j in 2:npred)
       yhat[j]=yhat[j-1]+yhat[j]
     # store
     predictions.addact[i,] <- yhat
}
```

Simple plots
```{r fig.height=6, fig.width=10}
offset=1
plotr=3
plotc=4
par(mfrow=c(plotr,plotc))
par('mar'=c(1,1,1,1))
for (i in 1:(plotr*plotc))
{
  ind=offset*plotr*plotc+i
  #plot(cgm[(n+(i-1)*timesteps):(n+timesteps*i)],type='l')
  plot(cgm[ind,],type='l',ylim=c(min(3,min(cgm[ind,])),max(15,max(cgm[ind,]))))
  points(x=(1+timesteps-npred):(timesteps),y=predictions[ind,],col='red', pch=20)
  
  # uncomment to plot imaginary bolus/activity predictions
  # points(x=(1+timesteps-npred):(timesteps),y=predictions.addbol[ind,],col='blue', pch=20)
  # points(x=(1+timesteps-npred):(timesteps),y=predictions.addact[ind,],col='green', pch=20)

}
```

Performance metrics - From Annie

```{r}
getPerformance = function(pred, val) {
    res = pred - val
    MAE = sum(abs(res))/length(val)
    RSS = sum(res^2)
    MSE = RSS/length(val)
    RMSE = sqrt(MSE)
    perf = data.frame(MAE, RSS, MSE, RMSE)
}
```

```{r}
print(getPerformance(predictions.raw,y_test))
```
Save the model
```{r}
# save_model_hdf5(model, paste("model",timesteps,"denoise.h5",sep='-'))
```

Code to load a saved model
```{r}
# load_model_hdf5('model-288-diff.h5', custom_objects = NULL, compile = TRUE)
```

